% Dedication
 \dedication{To my wife, Jinjuan, and our son, Charles.}%ZZZZ

% Acknowledgments
 \acknowledgments[6]{
  While this thesis represents my work over the past 5 years, this work would not exist without the contributions from many different people.

  I would like to thank the entirety of my thesis committee for guiding my research, and offering insights.
  In particular I would like to thank my chair, Professor Brendan Kochunas, whose thesis was the original inspiration for this work.
  I am thankful for his continued mentorship throughout my PhD studies, and his clear interest and dedication to helping this work progress.
  I am especially thankful to Professor Thomas Downar, my original advisor, who has been instrumental to the development of my studies.
  As an undergraduate, Professor Downar gave me the opportunity to work in his research group.
  Without this experience it is unlikely I would have pursued my PhD studies.
  I'd like to thank Professor Edward Larsen, whose passion for neutron transport has helped to keep me motivated throughout my studies.
  I owe a special thanks to Dr. Rodolfo Ferrer, whose work with the linear source served as a foundation to base parts of this work.

  The work of this thesis has relied heavily on the MPACT code;
    without the hard work of all the MPACT team developers, this work would certainly not have been possible for me.
  I am grateful for the opportunity to have worked on this code, alongside my colleagues, for the past 6 years.
  I would also like to express my gratitude to the \ac{DOE}, for funding the majority of my PhD studies,
    through the Nuclear Energy University Graduate fellowship, and through the CASL program.

  Finally, I would like to thank my family, friends, and colleagues.
  My parents, who inspired my passion for mathematics and computer science.
  To my wife, Jinjuan, who has helped to keep me motivated throughout my studies.
  And my friends and colleagues who have allowed me to vent my countless struggles and frustrations over the course of my studies.
 }
%  This command sets the width of the acknowledgments area as a fraction
%  of the total width of the text area.
 \acknowledgmentswidth{0.95} %ZZZZ